Full-core analysis of thorium-fueled Molten Salt Breeder Reactor using the SERPENT 2 Monte Carlo code

### 2017-08-13 full-core_msbr_model.tex
grammar

Zone II is Zone II 37\% salt by volume and each element has a fuel channel diameter of 
6.604cm. It is divided into two different zones: Zone II-A and Zone II-B. The 
graphite elements for Zone II-A are prismatic. Zone II-B elements are in the 
form of rectangular slats spaced far enough apart to provide the 0.37 fuel salt 
volume fraction. Fig.~\ref{fig:zone2B} additionally shows the 5.08cm-wide 
annular space between the core graphite and the radial reflector graphite. The 
annulus contains 100\% fuel salt and serves to reduce the damage flux at the 
internal surface of the graphite reflector blocks. The reactor Zone II-B 
graphite 5.08cm-thick slats and various length in width (average width is about 
26.67cm) are reconstructed as is in the model without any approximation. From 
the ORNL report \cite{robertson_conceptual_1971}, the suggested design of Zone 
II-B has 8 irregularly-shaped graphite elements every 45$^\circ$ as well as 
salt channels. These graphite elements were simplified into right-circular 
cylindrical shapes with central channels. This is the only simplification made 
to the \gls{MSBR} conceptual geometry in this work.
\FloatBarrier
\begin{figure}[htp!] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=\linewidth]{figure_2_4.png}
  \caption{Zone I (left) and Zone II-A (right) elements.}
  \label{fig:zone12A}
\end{figure}
\subsection{Material composition}
The fuel salt, the reactor graphite, and the modified HastelloyN are materials 
unique of the \gls{MSBR} and were created at \gls{ORNL}. The initial fuel salt loading composition is LiF-BeF$_2$-ThF$_4$-$^{233}$UF$_4$ (71.8-16-12-0.2 mole \%). The lithium in the molten salt fuel is a fully enriched $^{7}$Li because $^{6}$Li is a very strong neutron poison and becomes tritium upon neutron capture. For cross section data generation ENDF/B-VII was employed \cite{chadwick_endf/b-vii.0:_2006}. The specific temperature was fixed for each material to correctly model the Doppler-broadening of resonance peaks when Serpent generate problem-oriented nuclear data library.
%The isotope composition of each material at the initial state was described in detail in the MSBR conceptual design study \cite{robertson_conceptual_1971} and has been applied to Serpent model without any modification. %
\begin{figure}[htbp!] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=0.96\linewidth]{figure_2_5.png}
  \caption{Plan view of Zone II-B.}
  \label{fig:zone2B}
\end{figure}
\FloatBarrier
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and Analysis}
The reactor physics calculation was carried out using the methods described earlier. This section presents calculation results, such as the effective multiplication factor for whole core, neutron flux spectrum, and temperature reactivity coefficients. The normalized neutron flux distribution is calculated for the whole core using continuous-energy nuclear data which allows to obtaining very smooth distribution. The temperature coefficients for both fuel salt solution and reactor graphite are computed by comparing effective multiplication factors for two temperatures in the working range.

\subsection{Neutron spectrum}
Fig.~\ref{fig:spectrum} demonstrates the normalized neutron flux spectrum for the whole core in the energy range from $10^{-9}$ to $10$ MeV. The results show close fit with the MCNP simulation \cite{park_whole_2015}, especially in thermal energy range. 
\begin{figure}[h!] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=1.05\linewidth]{figure_3_1.png} 
  \caption{Neutron flux spectrum of \gls{MSBR} for MCNP6 and Serpent 2 model.}
  \label{fig:spectrum}
\end{figure}
It is important to obtain the epithermal and thermal spectrum to produce $^{233}$U from $^{232}$Th because radiative capture cross section of thorium monotonically decreasing from $10^{-10}$ MeV to $10^{-5}$ MeV. Hardening the spectrum tends to significantly increase resonance absorption in thorium and decreasing the absorptions in fissile and construction materials thus large amount of fissile material will be needed to make the reactor critical.

\subsection{Effective multiplication factor}
Table~\ref{tab:keff} shows the effective multiplication factor for both MCNP6 and Serpent 2 whole core models. The factor obtained using Serpent 2 is about 300 pcm lower than MCNP6 one, while standard deviations are 5 and 9 pcm, respectively. The discrepancy might occurred because the model used in Park \emph{et al.} has few simplification in Zone II geometry.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\captionsetup[table]{
  labelsep = newline,
  name = TABLE, 
  justification=justified,
  singlelinecheck=false,%%%%%%% a single line is centered by default
  labelsep=colon,%%%%%%
  skip = \medskipamount}
\begin{table}[h!]
%\centering
\caption{Effective multiplication factor of whole core model.}
\begin{tabular}{p{0.15\linewidth} p{0.3\linewidth} p{0.3\linewidth}} \toprule
      & Serpent2      & MCNP6 \cite{park_whole_2015}          
\\ \midrule
K$_{eff}$  & 1.00389$\pm$0.00005 & 1.00736$\pm$0.00009
\\
\bottomrule
\end{tabular}
  \label{tab:keff}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Temperature effect of reactivity}
Table~\ref{tab:tcoef} represents the quantitive analysis of temperature effects on reactivity. Uncertainty for each temperature coefficient also was calculated and shown in Table~\ref{tab:tcoef}. The main physical principle underlying the reactor temperature feedback is an expansion of matter when it is heated. When the fuel salt temperature increases, the density of the salt decreases, but at the same time, the total volume of fuel salt in the core remains constant because it is defined by the space for fuel bounded by the graphite. When the reactor graphite temperature grows, the density of graphite declines which also frees up space for fuel salt. To determine temperature coefficients the cross-section temperatures for fuel and moderator were changed from 900K to 1200K. Three different cases were considered:
\begin{enumerate}  
\item Temperature of fuel salt rising from 900K to 1200K.
\item Temperature of graphite rising from 900K to 1200K. 
\item Whole reactor temperature rising from 900K to 1200K.
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\captionsetup[table]{
  labelsep = newline,
  name = TABLE, 
  justification=justified,
  singlelinecheck=false,%%%%%%% a single line is centered by default
  labelsep=colon,%%%%%%
  skip = \medskipamount}
\begin{table}[h!]
%\centering
  \caption{Temperature coefficients of reactivity.}
\begin{tabular}{p{0.22\linewidth} p{0.22\linewidth} p{0.21\linewidth} p{0.15\linewidth}} \toprule
   Reactivity coefficient [pcm/K]  & Serpent2      & MCNP6 \cite{park_whole_2015}   & Reference \cite{robertson_conceptual_1971}      
\\ \midrule
Fuel salt        & $-3.70\pm0.016$ & $-3.20\pm0.05$ & $-3.22$ 
\\ \midrule
Moderator        & $+2.33\pm0.027$ & $-0.11\pm0.05$ & $+2.35$ 
\\ \midrule
Total            & $-1.57\pm0.033$ & $-3.21\pm0.04$ & $-0.87$ 
\\
\bottomrule
\end{tabular}
  \label{tab:tcoef}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
On the one hand, changes in fuel temperature cause only density variation, geometry keeps the same because fuel is in the form of liquid. On the other hand, when moderator heats up both the density and the geometry changes due to thermal expansion of the graphite blocks and the reflector. New graphite density was calculated using linear temperature expansion coefficient of reactor graphite which is 1.3$\times10^{-6}$1/K \cite{robertson_conceptual_1971}. Based on this information new geometry input, which takes into account graphite expansion, was created.

>>>>

Zone II is Zone II 37\% salt by volume and each element has a fuel channel diameter of 
6.604cm. It is divided into two different zones: Zone II-A and Zone II-B. The 
graphite elements for Zone II-A are prismatic. Zone II-B elements are 
rectangular slats spaced far enough apart to provide the 0.37 fuel salt 
volume fraction. Fig.~\ref{fig:zone2B} additionally shows the 5.08cm-wide 
annular space between the core graphite and the radial reflector graphite. The 
annulus contains 100\% fuel salt and serves to reduce the damage flux at the 
internal surface of the graphite reflector blocks. The reactor Zone II-B 
graphite 5.08cm-thick slats vary in the radial dimension (average width is 
26.67cm) but are reconstructed without any approximation. From 
the ORNL report \cite{robertson_conceptual_1971}, the suggested design of Zone 
II-B has 8 irregularly-shaped graphite elements every 45$^\circ$ as well as 
salt channels. These graphite elements were simplified into right-circular 
cylindrical shapes with central channels. This is the only simplification made 
to the \gls{MSBR} conceptual geometry in this work.
\FloatBarrier
\begin{figure}[htp!] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=\linewidth]{figure_2_4.png}
  \caption{Zone I (left) and Zone II-A (right) elements.}
  \label{fig:zone12A}
\end{figure}
\subsection{Material composition}
The fuel salt, the reactor graphite, and the modified HastelloyN are materials 
unique of the \gls{MSBR} and were created at \gls{ORNL}. The initial fuel salt 
loading composition is LiF-BeF$_2$-ThF$_4$-$^{233}$UF$_4$ (71.8-16-12-0.2 mole 
\%). The lithium in the molten salt fuel is a fully enriched $^{7}$Li because 
$^{6}$Li is a very strong neutron poison and becomes tritium upon neutron 
capture. For cross section generation, ENDF/B-VII was employed \cite{chadwick_endf/b-vii.0:_2006}. The specific temperature was fixed for each material to correctly model the Doppler-broadening of resonance peaks when Serpent generate problem-oriented nuclear data library.
%The isotope composition of each material at the initial state was described in detail in the MSBR conceptual design study \cite{robertson_conceptual_1971} and has been applied to Serpent model without any modification. %
\begin{figure}[htbp!] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=0.96\linewidth]{figure_2_5.png}
  \caption{Plan view that includes Zone I, II-A, and II-B elements.}
  \label{fig:zone2B}
\end{figure}
\FloatBarrier
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{R}
This section presents calculation results, such as the effective multiplication 
factor for whole core, neutron flux spectrum, and temperature reactivity 
coefficients. The normalized neutron flux distribution is calculated for the 
whole core using continuous-energy nuclear data.
The temperature coefficients for both fuel salt solution and reactor graphite are computed by comparing effective multiplication factors for two temperatures in the working range.

\subsection{Neutron spectrum}
Fig.~\ref{fig:spectrum} demonstrates the normalized neutron flux spectrum for the whole core in the energy range from $10^{-9}$ to $10$ MeV. The results show close fit with the MCNP simulation \cite{park_whole_2015}, especially in thermal energy range. 
\begin{figure}[h!] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[width=1.05\linewidth]{figure_3_1.png} 
  \caption{Neutron flux spectrum of \gls{MSBR} for MCNP6 and Serpent 2 model.}
  \label{fig:spectrum}
\end{figure}
It is important to obtain the epithermal and thermal spectrum to produce 
$^{233}$U from $^{232}$Th because radiative capture cross section of thorium 
monotonically decreases from $10^{-10}$ MeV to $10^{-5}$ MeV. Hardening the 
spectrum tends to significantly increase resonance absorption in thorium and 
decrease the absorptions in fissile and construction materials. Thus, 
a signficant amount fissile material will be needed to make the reactor critical.

\subsection{Effective multiplication factor}
Table~\ref{tab:keff} shows the effective multiplication factor for both MCNP6 
and Serpent 2 whole core models. The factor obtained using Serpent 2 is 300 pcm 
lower than that obtained by Park \emph{et al.} using MCNP6 \cite{park_whole_2015}. Standard 
deviations are 5 and 9 pcm, respectively. The discrepancy is likely due to 
simplificiations to the Zone II geometry model used in Park \emph{et al.}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\captionsetup[table]{
  labelsep = newline,
  name = TABLE, 
  justification=justified,
  singlelinecheck=false,%%%%%%% a single line is centered by default
  labelsep=colon,%%%%%%
  skip = \medskipamount}
\begin{table}[h!]
%\centering
\caption{Effective multiplication factor of whole core model.}
\begin{tabular}{p{0.15\linewidth} p{0.3\linewidth} p{0.3\linewidth}} \toprule
      & Serpent2      & MCNP6 \cite{park_whole_2015}          
\\ \midrule
K$_{eff}$  & 1.00389$\pm$0.00005 & 1.00736$\pm$0.00009
\\
\bottomrule
\end{tabular}
  \label{tab:keff}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Temperature effect of reactivity}
Table~\ref{tab:tcoef} shows temperature effects on reactivity calculated in 
this work as compared to both \cite{park_whole_2015} and 
\cite{robertson_conceptual_1971}. Uncertainty for each temperature coefficient 
also appears in Table~\ref{tab:tcoef}. The main physical principle underlying 
the reactor temperature feedback is an expansion of matter when it is heated. 
When the fuel salt temperature increases, the density of the salt decreases, 
but at the same time, the total volume of fuel salt in the core remains 
constant because it is bounded by the graphite. When the reactor graphite 
temperature grows, the density of graphite declines creating additional space 
for fuel salt. To determine temperature coefficients, the cross-section temperatures for fuel and moderator were changed from 900K to 1200K. Three different cases were considered:
\begin{enumerate}  
\item Temperature of fuel salt rising from 900K to 1200K.
\item Temperature of graphite rising from 900K to 1200K. 
\item Whole reactor temperature rising from 900K to 1200K.
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\captionsetup[table]{
  labelsep = newline,
  name = TABLE, 
  justification=justified,
  singlelinecheck=false,%%%%%%% a single line is centered by default
  labelsep=colon,%%%%%%
  skip = \medskipamount}
\begin{table}[h!]
%\centering
  \caption{Temperature coefficients of reactivity.}
\begin{tabular}{p{0.22\linewidth} p{0.22\linewidth} p{0.21\linewidth} p{0.15\linewidth}} \toprule
   Reactivity coefficient [pcm/K]  & Serpent2      & MCNP6 \cite{park_whole_2015}   & Reference \cite{robertson_conceptual_1971}      
\\ \midrule
Fuel salt        & $-3.70\pm0.016$ & $-3.20\pm0.05$ & $-3.22$ 
\\ \midrule
Moderator        & $+2.33\pm0.027$ & $-0.11\pm0.05$ & $+2.35$ 
\\ \midrule
Total            & $-1.57\pm0.033$ & $-3.21\pm0.04$ & $-0.87$ 
\\
\bottomrule
\end{tabular}
  \label{tab:tcoef}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the first case, changes in the fuel temperature only impact fuel density. In 
this case, the geometry is unchanged because fuel is a liquid. However, when 
the moderator heats up, both the density and the geometry change due to thermal 
expansion of the solid graphite blocks and reflector. Accordingly, the new 
graphite density was calculated using a linear temperature expansion 
coefficient of 1.3$\times10^{-6}$1/K \cite{robertson_conceptual_1971}. A new 
geometry input was created based on this information.
